### 2014-12-11 ms/ms.Rmd
Minor tweaks to user information, and manuscript

We (D.S. Falster and R.A. Duursma, with the help of D.R. Barneche, R.G. FitzJohn and A. Vårhammar) set out to create such a resource. We identified data sources from existing collaborations as well as a wide-ranging literature search, focussed on woody plants in field as well as controlled conditions. We constrained the search by including only data where a wide range in plant sizes was measured, as well as a set of minimum variables (plant height and leaf mass or area), and an emphasis on field-grown plants (nearly all data from plants grown in controlled conditions arose from existing collaborations). We do not claim to have done an exhaustive search of the literature, due to the very large number of potential studies that could be included. In each case we asked authors directly whether they would be willing to make their raw data files freely available. The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data. Moreover, we were invited to incorporate another compilation led by M.I. Ishihara and focusing on Japanese literature. As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies.

1. Our entire work flow is scripted, using the R computing environment. The database can be rebuilt in an instant from the raw data files, using a single command. When a single data contribution has been modified, the final master dataset is thus easily recreated.
2. We minimised the amount of code by requiring each study to conform to a common format (Figure 2).
3. We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`. These reports were sent to data contributors for verification, and thus allowed for easy identification of errors in the raw data or processing. Each report included plots for every pairwise combination of quantitative variables in a given study, site maps, and metadata. The updated reports are available on the database website.
4. We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data. Datasets as well as processing methods inevitably contain errors. With version control, we are able to trace these errors to the source.

@@

**Figure 1** Global distribution of studies included in BAAD. Symbols are sized relative to the number of individual plants in each dataset.

![baadworkflow](Figure2.png)
**Figure2** Workflow of BAAD. Each of the raw data files was processed in the same way, resulting in a single dataset ('BAAD') with a common format.

@@

### A. DATA SET IDENTITY
BAAD: a Biomass And Allometry Database for woody plants

@@

**2. Abstract**: Understanding how plants are plant constructed -- i.e. how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping energy-fluxes in the terrestrial biosphere. Allocation patterns can differ through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments. While a variety of models dealing with biomass allocation exist, we lack a synthetic understanding of the underlying processes. This is partly due to the lack of suitable datasets for validating and parameterising models. To that end, we present the Biomass And Allometry Database (BAAD) for woody plants. BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species. Most of these data are associated with an existing publication, however, raw data files were rarely made public at time of publication. Thus BAAD contains individual level data from different studies, transformed into standard units and variable names. The transformations were achieved using a common workflow for all raw data files. Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e. not indirectly via allometric equations (except in very large trees where biomass was estimated from detailed sub-sampling). We included both wild and artificially grown plants. The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues. We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, light, experimental treatment, vegetation type) where available. It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the world's vegetation.

@@

**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in BAAD, and to minimize outliers.

@@

BAAD contains the following elements:

>>>>

We (D.S. Falster and R.A. Duursma, with the help of D.R. Barneche, R.G. FitzJohn and A. Vårhammar) set out to create such a resource. We identified data sources from existing collaborations as well as a wide-ranging literature search, focussed on woody plants in field as well as controlled conditions. We constrained the search by including only data where a wide range in plant sizes was measured, as well as a set of minimum variables (plant height and leaf mass or area), and an emphasis on field-grown plants (nearly all data from plants grown in controlled conditions arose from existing collaborations). We do not claim to have done an exhaustive search of the literature, due to the very large number of potential studies that could be included. In each case we asked authors directly whether they would be willing to make their raw data files freely available. The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data. Moreover, we were invited to incorporate another compilation led by M.I. Ishihara and focusing on Japanese literature. As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies (Fig. 1).

@@

**Figure 1** Global distribution of studies included in the BAAD. Symbols are sized relative to the number of individual plants in each dataset.

1. Our entire work flow is scripted, using the R computing environment. The database can be rebuilt in an instant from the raw data files, using a single command. When a single data contribution has been modified, the final master dataset is thus easily recreated.
2. We minimised the amount of code by requiring each study to conform to a common format, with separate files for raw data, units of data, meta data, contact details, and citation (see examples in attached code). These files were then processed in a standardised way (Fig. 2).
3. We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`. These reports were sent to data contributors for verification, and thus allowed for easy identification of errors in the raw data or processing. Each report included plots for every pairwise combination of quantitative variables in a given study, site maps, and metadata. The updated reports are available on the database website.
4. We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data. Datasets as well as processing methods inevitably contain errors. With version control, we are able to trace these errors to the source, and review review the history of changes to any given dataset.

![baadworkflow](Figure2.png)
**Figure2** Workflow for building the BAAD. Data from each study is processed in the same way, using a standardised set of input files, resulting in a single dataset with a common format.

@@

### A. DATA SET IDENTITY
the BAAD: a Biomass And Allometry Database for woody plants

@@

**2. Abstract**: Understanding how plants are plant constructed -- i.e. how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping energy-fluxes in the terrestrial biosphere. Allocation patterns can differ through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments. While a variety of models dealing with biomass allocation exist, we lack a synthetic understanding of the underlying processes. This is partly due to the lack of suitable datasets for validating and parameterising models. To that end, we present the Biomass And Allometry Database (BAAD) for woody plants. The BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species. Most of these data are associated with an existing publication, however, raw data files were rarely made public at time of publication. Thus the BAAD contains individual level data from different studies, transformed into standard units and variable names. The transformations were achieved using a common workflow for all raw data files. Other features which distinguish the BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e. not indirectly via allometric equations (except in very large trees where biomass was estimated from detailed sub-sampling). We included both wild and artificially grown plants. The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues. We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, light, experimental treatment, vegetation type) where available. It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the world's vegetation.

@@

**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in the BAAD, and to minimize outliers.

@@

The BAAD contains the following elements:
